Sweat electrolytes and CFTR activity in the era of CFTR modulators

### 2025-12-10 Edith Zemanick/Sweat Variability/sweat_test_analysis.qmd
sweat test formatting
updated report per Angela's changes
updated sweat test figures

---
title: "Sweat electrolytes and CFTR activity in the era of CFTR modulators"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 10
    toc-float: true
    number-sections: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
editor: source
---

```{r setup}
#| include: false
library(knitr)
library(Hmisc)
library(lubridate)
library(readxl)
library(tidyverse)
library(arsenal)
library(pROC)
home_dir <- switch(Sys.info()["sysname"],
    "Darwin" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Edith Zemanick/Sweat Variability",
    "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF/Edith Zemanick/Sweat Variability",
    "Linux" = "/home/tim/OneDrive/Vigers/CF/Edith Zemanick/Sweat Variability"
)
github_dir <- switch(Sys.info()["sysname"],
    "Darwin" = "/Users/tim/Documents/GitHub",
    "Windows" = "C:/Users/Tim/Documents/GitHub",
    "Linux" = "/home/tim/Documents/GitHub"
)
source(paste0(github_dir, "/CF/genotype_class_severity.R"))
knitr::opts_knit$set(root.dir = home_dir)
```

@@

@@

@@

    numeric.stats = c("N", "Nmiss", "meansd", "medianq1q3", "iqr", "range")

@@

```{r results='asis'}
summary(
    tableby(
        Diagnosis ~
            Method +
            Sex +
            Age +
            fe(`Genotype Risk`) +
            fe(`Cl Group`) +
            Cl +
            Na +
            `Na:Cl` +
            K,
        data = df,
        simulate.p.value = T,
        control = mycontrols
    ),
    pfootnote = T
)
```

@@

```{r results='asis'}
summary(
    tableby(
        `Genotype Risk` ~
            Sex +
            Age +
            fe(`Cl Group`) +
            Cl +
            Na +
            `Na:Cl` +
            K,
        data = df[
            df$`Genotype Risk` != "Unknown" & df$Diagnosis == "Cystic Fibrosis",
        ],
        simulate.p.value = T,
        control = mycontrols
    ),
    pfootnote = T
)
```

@@

```{r results='asis'}
summary(
    tableby(
        `Cl Group` ~ Sex + Age + Cl + Na + `Na:Cl` + K,
        data = df,
        control = mycontrols
    ),
    pfootnote = T
)
```

@@

```{r}
#| warning: false
clp <- ggplot(df, aes(x = Diagnosis, y = Cl)) +
    geom_boxplot() +
    theme_bw()
clp
```

@@

```{r}
#| warning: false
clp <- ggplot(df, aes(x = `Genotype Risk`, y = Cl)) +
    geom_boxplot() +
    theme_bw()
clp
```

@@

```{r}
#| warning: false
nap <- ggplot(df, aes(x = Diagnosis, y = Na)) +
    geom_boxplot() +
    theme_bw()
nap
```

@@

```{r}
#| warning: false
nap <- ggplot(df, aes(x = `Genotype Risk`, y = Na)) +
    geom_boxplot() +
    theme_bw()
nap
```

@@

```{r}
#| warning: false
ratiop <- ggplot(df, aes(x = Diagnosis, y = `Na:Cl`)) +
    geom_boxplot() +
    geom_hline(yintercept = 1, color = "red") +
    geom_hline(yintercept = 1.25, color = "red") +
    theme_bw() +
    ylab("Na:Cl") +
    xlab("Modulator")
ratiop
ratiop <- ggplot(df, aes(x = Diagnosis, y = `Na:Cl`)) +
    geom_boxplot() +
    geom_hline(yintercept = 1, color = "red") +
    geom_hline(yintercept = 1.25, color = "red") +
    ylim(0, 2) +
    theme_bw() +
    ylab("Na:Cl") +
    xlab("Modulator")
ratiop
```

@@

```{r}
#| warning: false
ratiop <- ggplot(df, aes(x = `Genotype Risk`, y = `Na:Cl`)) +
    geom_boxplot() +
    geom_hline(yintercept = 1, color = "red") +
    geom_hline(yintercept = 1.25, color = "red") +
    theme_bw() +
    ylab("Na:Cl") +
    xlab("Modulator")
ratiop
ratiop <- ggplot(df, aes(x = `Genotype Risk`, y = `Na:Cl`)) +
    geom_boxplot() +
    geom_hline(yintercept = 1, color = "red") +
    geom_hline(yintercept = 1.25, color = "red") +
    ylim(0, 2) +
    theme_bw() +
    ylab("Na:Cl") +
    xlab("Modulator")
ratiop
```

@@

```{r}
#| warning: false
ratiop <- ggplot(df, aes(x = `Genotype Risk`, y = `Na:Cl`)) +
    geom_boxplot() +
    geom_hline(yintercept = 1, color = "red") +
    geom_hline(yintercept = 1.25, color = "red") +
    theme_bw() +
    ylab("Na:Cl") +
    xlab("Modulator")
ratiop
ratiop <- ggplot(df, aes(x = `Genotype Risk`, y = `Na:Cl`)) +
    geom_boxplot() +
    geom_hline(yintercept = 1, color = "red") +
    geom_hline(yintercept = 1.25, color = "red") +
    ylim(0, 2) +
    theme_bw() +
    ylab("Na:Cl") +
    xlab("Modulator")
ratiop
```

>>>>

---
title: "Sweat electrolytes and CFTR activity in the era of CFTR modulators"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 10
    toc-float: true
    number-sections: true
    code-fold: true
    embed-resources: true
    fig-cap-location: top
    page-layout: full
editor: source
---

```{r setup}
#| include: false
library(knitr)
library(Hmisc)
library(lubridate)
library(readxl)
library(tidyverse)
library(arsenal)
library(pROC)
library(patchwork)
home_dir <- switch(
  Sys.info()["sysname"],
  "Darwin" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Edith Zemanick/Sweat Variability",
  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF/Edith Zemanick/Sweat Variability",
  "Linux" = "/home/tim/OneDrive/Vigers/CF/Edith Zemanick/Sweat Variability"
)
github_dir <- switch(
  Sys.info()["sysname"],
  "Darwin" = "/Users/tim/Documents/GitHub",
  "Windows" = "C:/Users/Tim/Documents/GitHub",
  "Linux" = "/home/tim/Documents/GitHub"
)
source(paste0(github_dir, "/CF/genotype_class_severity.R"))
knitr::opts_knit$set(root.dir = home_dir)
```

@@

# Manual fixes per Angela's 11/21/25 email
chco$Diagnosis[chco$PatientID == 549] = "CFTR-related disorder"
chco$Diagnosis[chco$PatientID == 650] = "Cystic Fibrosis"
chco$Diagnosis[chco$PatientID == 653] = "CFTR-related disorder"
chco$Diagnosis[chco$PatientID == 693] = "Cystic Fibrosis"
chco$Diagnosis[chco$PatientID == 702] = "Cystic Fibrosis"
chco$Diagnosis[chco$PatientID == 721] = "Cystic Fibrosis"
chco$Diagnosis[chco$PatientID == 736] = "CFTR-related disorder"
chco[nrow(chco) + 1, ] <- NA
chco$PatientID[nrow(chco)] = 692
chco$Age[nrow(chco)] = (mdy("3/24/2014") - mdy("9/26/13")) / 365.25
chco$Diagnosis[nrow(chco)] = "CRMS"
chco$Na_Left[nrow(chco)] = 24.1
chco$Na_Right[nrow(chco)] = 21.1
chco$Cl[nrow(chco)] = 30
chco$Cl_Right[nrow(chco)] = 30
chco[nrow(chco) + 1, ] <- NA
chco$PatientID[nrow(chco)] = 724
chco$Age[nrow(chco)] = (mdy("5/08/2015") - mdy("1/18/09")) / 365.25
chco$Diagnosis[nrow(chco)] = "CFTR-related disorder"
chco$Na_Left[nrow(chco)] = 72.4
chco$Na_Right[nrow(chco)] = 72.4
chco$Cl_Left[nrow(chco)] = 64
chco$Cl_Right[nrow(chco)] = 59

@@

# Age groups
df$`Age Group` = cut(
  df$Age,
  breaks = c(-Inf, 18, Inf),
  labels = c("< 18", ">= 18"),
  right = FALSE
# Specific genotypes
df$Genotypes1[
  !df$Genotypes1 %in%
    c(
      "F508del",
      "G542X",
      "G551D",
      "W1282X",
      "N1303K",
      "R117H"
] = "Other"
df$Genotypes2[
  !df$Genotypes2 %in%
    c(
      "F508del",
      "G542X",
      "G551D",
      "W1282X",
      "N1303K",
      "R117H"
] = "Other"
df$Genotype = paste0(df$Genotypes1, "/", df$Genotypes2)
df$Genotype = factor(
  df$Genotype,
    "F508del/F508del",
    "F508del/G542X",
    "F508del/G551D",
    "F508del/N1303K",
    "F508del/Other",
    "F508del/R117H",
    "F508del/W1282X",
    "G542X/G551D",
    "G542X/N1303K",
    "G542X/Other",
    "G542X/R117H",
    "G551D/Other",
    "G551D/R117H",
    "N1303K/G542X",
    "N1303K/N1303K",
    "N1303K/Other",
    "Other/F508del",
    "Other/G542X",
    "Other/Other",
    "Other/W1282X",
    "R117H/F508del",
    "R117H/Other",
    "W1282X/N1303K"
    "F508del/F508del",
    "F508del/G542X",
    "F508del/G551D",
    "F508del/N1303K",
    "F508del/Other",
    "F508del/R117H",
    "F508del/W1282X",
    "Other",
    "Other",
    "Other",
    "Other",
    "Other",
    "Other",
    "Other",
    "Other",
    "Other",
    "F508del/Other",
    "Other",
    "Other",
    "Other",
    "F508del/R117H",
    "Other",
    "Other"
# Order factor variables
df$Diagnosis = factor(
  levels = c("Non-CF", "CFTR-RD or CRMS", "Cystic Fibrosis")
df$`Genotype Risk` = factor(
  df$`Genotype Risk`,
  levels = c("Not CF", "Mild", "Severe", "Unknown")

@@

  numeric.stats = c("N", "Nmiss", "meansd", "medianq1q3", "iqr", "range"),
  simulate.p.value = TRUE

@@

```{r results='asis'}
summary(
  tableby(
    Diagnosis ~
      Method +
      Sex +
      Age +
      fe(`Age Group`) +
      fe(Genotype) +
      fe(`Genotype Risk`) +
      fe(`Cl Group`) +
      Cl +
      Na +
      `Na:Cl` +
      K,
    data = df,
    simulate.p.value = T,
    control = mycontrols
  ),
  pfootnote = T
)
```

@@

```{r results='asis'}
summary(
  tableby(
    `Genotype Risk` ~
      Sex +
      Age +
      fe(`Age Group`) +
      fe(Genotype) +
      fe(`Cl Group`) +
      Cl +
      Na +
      `Na:Cl` +
      K,
    data = df[
      df$`Genotype Risk` != "Unknown" & df$Diagnosis == "Cystic Fibrosis",
    ],
    simulate.p.value = T,
    control = mycontrols
  ),
  pfootnote = T
)
```

@@

```{r results='asis'}
summary(
  tableby(
    `Cl Group` ~ Sex +
      Age +
      fe(`Age Group`) +
      fe(Genotype) +
      Cl +
      Na +
      `Na:Cl` +
      K,
    data = df,
    control = mycontrols
  ),
  pfootnote = T
)
```

@@

```{r}
#| warning: false
group_n <- df |> group_by(Diagnosis) |> summarise(n = n()) |> mutate(pos = 160)
clp_d <- ggplot(df, aes(x = Diagnosis, y = Cl)) +
  geom_boxplot() +
  geom_label(
    data = group_n,
    aes(x = Diagnosis, y = pos, label = paste("n =", n))
  ) +
  theme_bw()
clp_d
```

@@

```{r}
#| warning: false
geno_n <- df |>
  group_by(`Genotype Risk`) |>
  summarise(n = n()) |>
  mutate(pos = 160)
clp_g <- ggplot(df, aes(x = `Genotype Risk`, y = Cl)) +
  geom_boxplot() +
  geom_label(
    data = geno_n,
    aes(x = `Genotype Risk`, y = pos, label = paste("n =", n))
  ) +
  theme_bw()
clp_g
```

@@

```{r}
#| warning: false
nap_d <- ggplot(df, aes(x = Diagnosis, y = Na)) +
  geom_boxplot() +
  geom_label(
    data = group_n,
    aes(x = Diagnosis, y = pos, label = paste("n =", n))
  ) +
  theme_bw()
nap_d
```

@@

```{r}
#| warning: false
nap_g <- ggplot(df, aes(x = `Genotype Risk`, y = Na)) +
  geom_boxplot() +
  geom_label(
    data = geno_n,
    aes(x = `Genotype Risk`, y = pos, label = paste("n =", n))
  ) +
  theme_bw()
nap_g
```

@@

```{r}
#| warning: false
group_n$pos = 4
ratiop_d <- ggplot(df, aes(x = Diagnosis, y = `Na:Cl`)) +
  geom_boxplot() +
  geom_label(
    data = group_n,
    aes(x = Diagnosis, y = pos, label = paste("n =", n))
  ) +
  geom_hline(yintercept = 1, color = "red") +
  geom_hline(yintercept = 1.25, color = "red") +
  theme_bw() +
  ylab("Na:Cl")
ratiop_d
```

@@

```{r}
#| warning: false
geno_n$pos = 4
ratiop_g <- ggplot(df, aes(x = `Genotype Risk`, y = `Na:Cl`)) +
  geom_boxplot() +
  geom_label(
    data = geno_n,
    aes(x = `Genotype Risk`, y = pos, label = paste("n =", n))
  ) +
  geom_hline(yintercept = 1, color = "red") +
  geom_hline(yintercept = 1.25, color = "red") +
  theme_bw() +
  ylab("Na:Cl")
ratiop_g
```

```{r}
p1 = (clp_d + ratiop_d) / (clp_g + ratiop_g)
p1
ggsave(
  filename = "./Dissemination/Figures/fig1.png",
  plot = p1,
  height = 9,
  width = 12,
  units = "in",
  dpi = 600
)
```

@@

```{r}
#| warning: false
geno_n$pos = 4
ratiop_g <- ggplot(df, aes(x = `Genotype Risk`, y = `Na:Cl`)) +
  geom_boxplot() +
  geom_label(
    data = geno_n,
    aes(x = `Genotype Risk`, y = pos, label = paste("n =", n))
  ) +
  geom_hline(yintercept = 1, color = "red") +
  geom_hline(yintercept = 1.25, color = "red") +
  theme_bw() +
  ylab("Na:Cl")
ratiop_g
```

```{r}
p2 = (clp_d + ratiop_d) / (clp_g + ratiop_g)
p2
ggsave(
  filename = "./Dissemination/Figures/fig2.png",
  plot = p2,
  height = 9,
  width = 12,
  units = "in",
  dpi = 600
)
```
